### 2015-07-26 ms-thesis/ver2/hmm/proof2.tex

There is a little complication for the start-up formula for $\beta$
recursion. In this case it would not be possible to start with $\beta(q_{T+1})$
because that state does not exist in our system. In that case, could we start
with $\beta(q_T)$ equaling vector of ones?

\begin{proof}
From \eqref{py} we can use the formula for $p(y)$ at time $T$ to confirm this.

\begin{eqnarray*}
p(y) &=& \sum_{q_T} \alpha(q_T)\beta(q_T) \\
&=& \sum_{i} \alpha(q_T^i=1)\beta(q_T^i=1) 
\end{eqnarray*}

Since we assumed $\beta(q_T)$ would be vector of ones, $\beta(q_T^i=1)$ will
always be 1, hence it can be dropped

\begin{eqnarray*}
&=& \sum_{i} \alpha(q_T^i=1)\\
&=& \sum_{i} p(y_0,...,y_T,q_T^i=1)\\
&=& p(y_0,...,y_T)\\
&=& p(y)
\end{eqnarray*}

which verifies our assumption.
\end{proof}